Transparency of COVID-19-related research in medical journals

### 2022-03-08 manuscript_covid_transparency.qmd
The first part of the results

The total number of COVID-19-related papers (open access and non-open access) was 1038, of which full texts of 657 papers (63.3%) were accessible via the EPMC. Of those, 337 were published in 2020 and 320 in 2021. However, as we could not retrieve seven full texts from the EPMC database due to technical issues, our final sample included 650 full-text articles (Supplementary figure 1).

These articles were published in 59 journals of which the top five were British Dental Journal (n=69), Journal of Dental Education (n=59), British Journal of Oral and Maxillofacial Surgery (n=51), Oral Diseases (n=47), and BMC Oral Health (n=40). The mean and the median of the number of citations to these articles were 7.7 (standard deviation (SD) = 43.15) and 2 (interquartile range (IQR) = 5). The highest cited article with 854 citations was an article published in the International Journal of Oral Sciences in 2020 (20).

About three-fourths of the articles (n=478) had a statement to disclose COI (73.5%, 95% CI: 70.0%--76.8%). Articles published in 2021 had a higher rate of COI disclosure (79.6% vs 67.8%, P \< 0.001). About four-tenths of the references had a funding statement (n=261) (40.2%, 95% CI: 36.5%--43.0%). Articles in 2021 had a funding statement more frequently than 2020 articles (49.1% vs 31.6%, P \< 0.001). Less than 1 in 20 of the articles (n=27) were registered beforehand (4.2%, 95% CI: 2.9%--6.0%) and the rate was six times higher in 2021 compared with 2020 (7.2% vs 1.1%, P \< 0.001). Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives. At first, we got four papers with shared data and none with shared code. After checking manually, we omitted four false positives from the shared data papers. One of them used the journal's template: "The data that support the findings of this study are openly available in \[repository name\] at \[DOI\]" (21). Hence finally, one paper had shared data (0.2%). This article was published in BMC Oral Health in 2021 (22).

A total of 117 articles were published in journals with no impact factor. Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively). Detailed information is illustrated in Table 1.

>>>>

The total number of COVID-19-related papers (open access and non-open access) was 92752, of which full texts of 73042 papers (78.7%) were accessible via the EPMC. Of those, 33137 were published in 2020 and 39905 in 2021. However, as we could not retrieve 43 full texts from the EPMC database due to technical issues, our final sample included 72999 full-text articles (Appendix 3).

These articles were published in 4892 journals of which the top five were PLoS ONE (n=6825), International Journal of Environmental Research and Public Health (n=2686), Medicine (n=2318), BMJ Open (n=1926), and eLife (n=942). The mean and the median of the number of citations to these articles were 3.1 (standard deviation (SD) = 21.76) and 0 (interquartile range (IQR) = 2). The highest cited article with 2915 citations was an article published in the Cell in 2020 [@Walls2020].

More than nine-tenth of the articles (n=67009) had a statement to disclose COI (91.79%, 95% CI: 91.59%--91.99%). Articles published in 2021 had a higher rate of COI disclosure (93.2% vs 90.1	%, P \< 0.001). More than eight-tenths of the references had a funding statement (n=60530) (82.92%, 95% CI: 82.64%--83.19%). Articles in 2021 had a funding statement more frequently than 2020 articles (83.9% vs 81.7%, P \< 0.001). Less than 1 in 6 of the articles (n=11457) were registered beforehand (15.69%, 95% CI: 15.43%--15.96%) and the rate was higher in 2021 compared with 2020 (16.2% vs 15.0%, P \< 0.001).

Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives. Less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: %--%). Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P \< 0.001). Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: %--%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).

A total of 51601 articles were published in journals with no impact factor. Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively). Detailed information is illustrated in Table 1.
